---
title: "Monty Hall Problem"
author: "Mathias Steilen"
date: "15/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Simulation

The game show host leads you to a wall with three closed doors. Behind one of the doors is the car of your dreams, and behind each of the other two is a goat. The three doors all have even chances of hiding the car.

The host, a trustworthy person who knows precisely what is behind each of the three doors, explains how the game will work. First, you will choose a door without opening it. The host will then open one of the two remaining doors to reveal a goat. When this has been done, you will be given the opportunity to switch doors. You will win whatever is behind the door you choose at this stage of the game.

Do you stick to your initial choice or do you switch doors?

Do you raise your chance of winning the car by switching doors?

***

```{r, include=FALSE}
rm(list = ls())
```

Define: 

- 1: Door 1
- 2: Door 2
- 3: Door 3
- C: Car
- G: Goat

First simulation: Not switching doors, sticking to the initial choice:

```{r, echo=FALSE}

#################################
## Simulation 1: Not switching ##
#################################

# params
simulations = 10000
outcomes <- c("C", "G", "G")
resmat <- matrix(ncol = 7, nrow = simulations)
colnames(resmat) <- c(1, 2, 3, "Initial Choice", 
                      "Revealed Door", "Subsequent choice", "Win")
options <- c(1:3)

# Simulation
for (i in 1:simulations){
  
  # host sets up the doors: 1 Car and 2 Goats
  resmat[i,1:3] <- sample(x = outcomes, size = 3, replace = FALSE)  

  # guest makes a random choice
  resmat[i,4] <- sample(x = options, size = 1, replace = FALSE)
  
  # host reveals a goat and gives choice between chosen closed
  # and unchosen closed door
  doors <- resmat[i,1:3]
  choice <- as.numeric(resmat[i,4])
  
  if (resmat[i,choice] == "C"){ # situation where guest chose car
    
    resmat[i,5] <- sample(x = options[! options %in% c(choice)], size = 1, 
                          replace = TRUE)
    
  } else { # situation where guest chose goat
    
    resmat[i,5] <- options[! options %in% c(choice,
                                          which(resmat[i,1:3] == "C")[[1]])]
    
  }
  
  # Offer to switch DECLINED
  resmat[i,6] = resmat[i,4]
  
  #Win or no win
  if ( resmat[i,6] == which(resmat[i,1:3] == "C")[[1]] ){
    
    resmat[i,7] = 1 # correct choice: win
    
  } else {
    
    resmat[i,7] = 0 # wrong choice: lose
    
  }
}

head(as.data.frame(resmat), 5)

```

```{r, fig.cap="Monty Hall Dilemma: Simulation - visualised", echo=FALSE}
###########
## PLOTS ##
###########

plotmat <- matrix(ncol = 2, nrow = simulations)
plotmat[,1] <- (1:simulations)
plotmat[,2] <- cumsum(resmat[,7])/(1:simulations)


plotnosw <- function(){
  plot(x = plotmat[,1], y = plotmat[,2], ylim = c(0,1), type = "l",
       ylab = "Win Probability",
       xlab = "Games",
       main = "Monty Hall Dilemma: Not Switching")
  abline(h = 1/3, col = "red", lty = "dashed")
  grid()
}

plotnosw()
```

Second simulation: Switching doors, not sticking to the initial choice:

```{r, echo=FALSE}

#############################
## Simulation 2: Switching ##
#############################

# params
resmat <- matrix(ncol = 7, nrow = simulations)
colnames(resmat) <- c(1, 2, 3, "Initial Choice", 
                      "Revealed Door", "Subsequent choice", "Win")

# Simulation
for (i in 1:simulations){
  
  # host sets up the doors: 1 Car and 2 Goats
  resmat[i,1:3] <- sample(x = outcomes, size = 3, replace = FALSE)  

  # guest makes a random choice
  resmat[i,4] <- sample(x = options, size = 1, replace = FALSE)
  
  # host reveals a goat and gives choice between chosen closed
  # and unchosen closed door
  doors <- resmat[i,1:3]
  choice <- as.numeric(resmat[i,4])
  
  if (resmat[i,choice] == "C"){ # situation where guest chose car
    
    resmat[i,5] <- sample(x = options[! options %in% c(choice)], size = 1, 
                          replace = TRUE)
    
  } else { # situation where guest chose goat
    
    resmat[i,5] <- options[! options %in% c(choice,
                                          which(resmat[i,1:3] == "C")[[1]])]
    
  }
  
  # Offer to switch ACCEPTED
  resmat[i,6] <- options[! options %in% c(choice, resmat[i,5])]
  
  #Win or no win
  if ( resmat[i,6] == which(resmat[i,1:3] == "C")[[1]] ){
    
    resmat[i,7] = 1 # correct choice: win
    
  } else {
    
    resmat[i,7] = 0 # wrong choice: lose
    
  }
}

head(as.data.frame(resmat))
```

```{r, fig.cap="Monty Hall Dilemma: Simulation - visualised", echo=FALSE}
###########
## PLOTS ##
###########

plotmat <- matrix(ncol = 2, nrow = simulations)
plotmat[,1] <- (1:simulations)
plotmat[,2] <- cumsum(resmat[,7])/(1:simulations)


plotsw <- function(){
  plot(x = plotmat[,1], y = plotmat[,2], ylim = c(0,1), type = "l",
       ylab = "Win Probability",
       xlab = "Games",
       main = "Monty Hall Dilemma: Switching")
  abline(h = 2/3, col = "red", lty = "dashed")
  grid()
}

plotsw()
```
